---
title: "Project3"
author: "Zhijun Liu"
date: "7/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DT)
library(tidyverse)
library(readr)
library(ggplot2)
library(GGally)
library(WVPlots)
library(pca3d)
library(tree)
```

# 1. Info

## description of data

[Attribute Information:](http://archive.ics.uci.edu/ml/datasets/Heart+failure+clinical+records)

Thirteen (13) clinical features:

- age: age of the patient (years)
- anaemia: decrease of red blood cells or hemoglobin (boolean)
- high blood pressure: if the patient has hypertension (boolean)
- creatinine phosphokinase (CPK): level of the CPK enzyme in the blood (mcg/L)
- diabetes: if the patient has diabetes (boolean)
- ejection fraction: percentage of blood leaving the heart at each contraction (percentage)
- platelets: platelets in the blood (kiloplatelets/mL)
- sex: woman or man (binary)
- serum creatinine: level of serum creatinine in the blood (mg/dL)
- serum sodium: level of serum sodium in the blood (mEq/L)
- smoking: if the patient smokes or not (boolean)
- time: follow-up period (days)
- [target] death event: if the patient deceased during the follow-up period (boolean)

## description of app


# 5. Data
```{r}
# read in data
data <- read_csv("heart_failure_clinical_records_dataset.csv") %>%
  rename( "CPK"=creatinine_phosphokinase,
          "EF" = ejection_fraction,
          "HT" = high_blood_pressure,
          "creatinine" = serum_creatinine,
          "sodium" = serum_sodium,
          "death" = DEATH_EVENT) %>%
  mutate(survive = 1-death) %>% select (-"death")


factorData <- data

factorData$anaemia <- as.factor(factorData$anaemia)
factorData$diabetes <- as.factor(factorData$diabetes)
factorData$HT <- as.factor(factorData$HT)
factorData$sex <- as.factor(factorData$sex)
factorData$smoking <- as.factor(factorData$smoking)
factorData$survive <- as.factor(factorData$survive)

datatable(data)
```

# 2. EDA
## categorical
### contigency table for select categorical variables
```{r}
table(data$anaemia,data$survive)
table(data$diabetes,data$survive)
table(data$HT,data$survive)
table(data$sex,data$survive)
table(data$smoking,data$survive)
```

### barchat
```{r}
# barchat for each 
ggplot(data = factorData) + 
  geom_bar(aes(x=anaemia,fill=survive),alpha=0.8) +
  scale_fill_brewer(palette="Set1")+
  labs(title='title')+ 
  theme_minimal()
```

## numeric
### statistic summaies for some numeric variable 
*default: all numeric elements*
```{r}
data_numeric_summary <- data %>% select(age,CPK,EF,platelets,creatinine,sodium,time)

apply(X=data_numeric_summary,2,FUN=function(x){summary(x)})
```

### correlation analysis for some elements
*default: all elements expect response*
```{r}

ggcorr(factorData,label = T, label_size = 3, angle= -30, label_round = 2,size=3)+
  ggplot2::labs(title = "Correlation Analysis")
```

### scatter plot for some numeric elements
*default: all numeric elements*
```{r}
data_numeric_scatter <- factorData %>% select(age,CPK,EF,platelets,creatinine,sodium,time,survive)


PairPlot(data_numeric_scatter, 
         colnames(data_numeric_scatter)[1:7], 
         title="Title", 
         group_var = "survive",
         alpha=0.5，
         palette="Set1")
```

### box plot for numeric element
*default: age*
```{r}
data_numeric_box <- data_numeric_scatter
ggplot(data = data_numeric_box,aes(x=survive,y=age,color=survive))+
  geom_violin(trim=FALSE,alpha=0.4)+
  geom_point(alpha=0.5,position = "jitter")+
  scale_colour_brewer(palette="Set1")

# CPK
# EF
# platelets
# creatinine
# sodium
# time

```

# 3. PCA

## Introduction
Principal components analysis (PCA) is a dimension reduction technique that is widely used in multivariate statistics. The objective is to condense the information that is present in the original set of variables via linear combinations of the variables while losing as little information as possible. 

Typically, the number of linear transformations is much smaller than the number of original variables; hence the reduction in the dimensionality of the data. This can be useful in different ways, such as providing better visualization and computational advantages. PCA also decorrelates the data, that is, PCA produces linear combinations of the variables that are mutually uncorrelated.

Goal: Obtain a linear combination of the variables that accounts for the largest amount of variability (assume 0 mean for each predictor).

$$ z_{i1} = \phi_{11}x_{i1}+ \phi_{21}x_{i2}+...+ \phi_{p1}x_{ip}$$
Constraint: 

- $\Sigma^p_{j=1}\phi^2_{j1}=1$.

- $\mathop{max}\limits_{\phi's}\frac{1}{n}\Sigma^n_{i=1}z^2_{i1}$.

- PCs must be mutually exclusive.

## analysis
```{r}
data_numeric_pc <- data_numeric_summary
pc.out <- prcomp(data_numeric_pc,scale = T)
# referrence
summary(pc.out)
screeplot(pc.out,type="lines",col="#990000",lwd=2)
# rotation
round(pc.out$rotation[,1:3],3)

pca3d(pc.out,components = c(1,2,3),group=data$survive, legend="right")
# snapshotPCA3d(file="second_plot.png")
pca2d(pc.out,components = c(1,2),group=data$survive,biplot = T, biplot.vars = 7, legend="right")
```


# 4. Modelling

## Tree model 

### Introduction
Classification tree if goal is to classify (predict) group membership.

For a binary response, within a given node (p=P(correct classification))

$$ Deviance: -2plog(p)-2(1-p)log(1-p)$$

For all possible splits, minimize this (weighted appropriately for node size).

We need to pruned back using classification error rate to prevent overfit, which might increase bias but decrease variance, hopefully improving prediction.

### Analysis
```{r}

# fit the tree model
treeFit <- tree(survive ~ . ,data=factorData,split="deviance")
# result of tree model
summary(treeFit)
# dendrogram
plot(treeFit)
text(treeFit, pretty = 0, cex = 0.6)
# giveing the referrence result for choosing tree size
pruneFit <- cv.tree(treeFit, FUN = prune.misclass)
pruneFit
# giveing the referrence plot for choosing tree size
plot(pruneFit$size,pruneFit$dev,type="b")
pruneFitFinal <- prune.misclass(treeFit,best=8) # choose the tree size in best option
# user predict to change the variable value (need to split the categorical data and numeric data)
# predict(pruneFitFinal, newdata = data.frame(time=10, creatinine=2),type="class")
```

## Logistic Regression
*default variable: time + creatinine*

### Introduction

Logistic regression is a statistical model that in its basic form uses a logistic function to model a binary dependent variable. 

Basic Logistic Regression models survive probability using logistic function, where $\boldsymbol{\beta} = (\beta_1,...\beta_p)^T,\boldsymbol{x}=(x_1,...x_p)^T$

$$P(survive=1|\boldsymbol{x})=\frac{e^{\alpha+\boldsymbol{\beta}^T\boldsymbol{x}}}{1+e^{\alpha+\boldsymbol{\beta}^T\boldsymbol{x}}}$$

$$log\left[\frac{P(survive=1|\boldsymbol{x})}{P(survive=0|\boldsymbol{x})}\right]=\alpha+\boldsymbol{\beta}^T\boldsymbol{x}$$

### Analysis

```{r}
glmFit <- glm(survive ~ ., data=factorData,family = "binomial")
glmFit
summary(glmFit)
# predict
# exp(predict(glmFit, newdata = data.frame(time=10, creatinine=2),type="link"))
```

